\documentclass[a4paper,oneside,12pt]{book}

\newcommand{\thesistitle}{French News Sentiment Analysis on COVID-19} % Your thesis title, this is used in the title and abstract
\newcommand{\degree}{B. A. (Mod.) Computer Science} % Your degree name, this is used in the title page and abstract
\newcommand{\typeofthesis}{dissertation} % dissertation, Final Year Project, report, etc.
\newcommand{\authorname}{Samuel Petit}

\newcommand{\keywords}{this, that, more} % Keywords for your thesis
\newcommand{\school}{\href{http://www.scss.tcd.ie}{School of Computer Science and Statistics}} % Your school's name and URL, this is used in the title page
\newcommand{\schoolAddress}{O’Reilly Institute, Trinity College, Dublin 2, Ireland} % Your school's name and URL, this is used in the title page

%% Comment out the next line if you don't want a department to appear
\newcommand{\trinity}{\href{https://www.tcd.ie/}{Trinity College}} 
\newcommand{\TRINITY}{\href{https://www.tcd.ie/}{TRINITY COLLEGE}} 

\AtBeginDocument{
\hypersetup{pdftitle=\thesistitle} % Set the PDF's title to your title
\hypersetup{pdfauthor=\authorname} % Set the PDF's author to your name
\hypersetup{pdfkeywords=\keywords} % Set the PDF's keywords to your keywords
\hypersetup{pdfsubject=\degree} % Set the PDF's keywords to your keywords
}

%% Language and font encodings
\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{csvsimple}
\usepackage{fancyvrb}
\usepackage{longtable}

%% Bibliographical stuff
\usepackage{natbib}

%% Document size
% include showframe as an option if you want to see the boxes
\usepackage[a4paper,top=2.54cm,bottom=2.54cm,left=2.54cm,right=2.54cm,headheight=16pt]{geometry}
\setlength{\marginparwidth}{2cm}
%% Useful packages
\usepackage{amsmath}
\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography
\usepackage[pdftex]{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=black]{hyperref}
\usepackage{caption} % if no caption, no colon
\usepackage{sfmath} %use sans-serif in the maths sections too
\usepackage[parfill]{parskip}    % Begin paragraphs with an empty line rather than an indent
\usepackage{setspace} % to permit one-and-a-half or double spacing
\usepackage{enumerate} % fancy enumerations like (i) (ii) or (a) (b) and suchlike
\usepackage{booktabs} % To thicken table lines
\usepackage{fancyhdr}
\usepackage{lscape}
\usepackage{float}

\pagestyle{plain} % Embrace simplicity!

\renewcommand{\familydefault}{\sfdefault} %use the sans-serif font as default

\renewcommand{\theequation}{\arabic{equation}} %% use continuous equation numbers

% Setup auto CSV to table commands 
\makeatletter
\csvset{
  autotabularcenter/.style={
    file=#1,
    after head=\csv@pretable\begin{tabular}{|*{\csv@columncount}{c|}}\csv@tablehead,
    table head=\hline\csvlinetotablerow\\\hline,
    late after line=\\,
    table foot=\\\hline,
    late after last line=\csv@tablefoot\end{tabular}\csv@posttable,
    command=\csvlinetotablerow},
  autobooktabularcenter/.style={
    file=#1,
    after head=\csv@pretable\begin{tabular}{*{\csv@columncount}{c}}\csv@tablehead,
    table head=\toprule\csvlinetotablerow\\\midrule,
    late after line=\\,
    table foot=\\\bottomrule,
    late after last line=\csv@tablefoot\end{tabular}\csv@posttable,
    command=\csvlinetotablerow},
}
\makeatother
\newcommand{\csvautotabularcenter}[2][]{\csvloop{autotabularcenter={#2},#1}}
\newcommand{\csvautobooktabularcenter}[2][]{\csvloop{autobooktabularcenter={#2},#1}}

%% Format Chapter headings appropriately
\usepackage{titlesec}
\titleformat{\chapter}[hang] 
{\normalfont\huge\bfseries}{\thechapter}{1cm}{} 

\title{\thesistitle}
\author{\authorname}

\frontmatter
\begin{document}
\input{title/title.tex}
\pagenumbering{roman}
\section*{\Huge{Declaration}}
\vspace{1cm}
I hereby declare that this \typeofthesis\ is entirely my own work and that it has not been submitted as an exercise for a degree at this or any other university.

\vspace{1cm}
I have read and I understand the plagiarism provisions in the General Regulations of the University Calendar for the current year, found at \url{http://www.tcd.ie/calendar}.
\vspace{1cm}

I have also completed the Online Tutorial on avoiding plagiarism `Ready Steady Write', located at \url{http://tcd-ie.libguides.com/plagiarism/ready-steady-write}.
\vspace{3cm}

Signed:~\rule{5cm}{0.3pt}\hfill Date:~\rule{5cm}{0.3pt}

\chapter*{Abstract}

News organisations play a major role in modern society and have been subject to many studies which rely on news articles and computer systems to extract relationships that are not always obvious. Given the importance of the English language in the field of research and computer science, less attention has been given to other languages such as French. Moreover, COVID-19 has recently impacted many aspects of society, putting news articles in a very important position. This thesis sets out to examine which variables have an impact on the language used in French news when writing articles about COVID-19.

A corpus of 305,454 French news articles, consisting of over 167 million words was collected from 226 sources starting from the beginning of the COVID-19 pandemic up until the day this study took place. The date range for data collection started in January 2020 and ended in March 2021. Three specialised dictionaries were built in order to perform topic specific analyses. These lexica all relate to key topics of the COVID-19 pandemic and target the sentiments of: Death, Vaccine and Virus. Through the use of sentiment analysis and natural language processing techniques the sentiment of the corpus was computed.

The sentiment proxies obtained from the corpus were used to identify underlying relationships affecting the amount of sentiment that can be found within French news articles. This was done through the use of Vector autoregression models which enables to identify and quantify causality between multiple variables.

\newpage
\onehalfspacing\raggedright %\raggedright turns off justification and hypenation

\section*{\Huge{Acknowledgements}}

I would like to thank my supervisor Professor Khurshid Ahmad for introducing me to this topic and guiding me throughout this project. I started with very little knowledge on many of the topics covered in this thesis and I have no doubt that your supervision and support throughout the year was a key factor in what I have learned as well as completing this project.

I would also like to thank my parents for pushing me to study abroad at Trinity, doing so changed my life in many ways that I could have never imagined. Throughout my time as a Trinity student I have learned so much and made many invaluable connections.

\tableofcontents

\mainmatter
\listoffigures
\listoftables
\input{introduction/introduction.tex}
\input{lit_review/lit_review}
\input{method/method}
\input{evaluation/evaluation.tex}
\input{conclusion/conclusion.tex}
\bibliographystyle{agsm}
\bibliography{bibs/sample}
\end{document}
