\chapter{Introduction}\label{Introduction}

This study focuses on extracting sentiment proxies from the French news media in order to understand news media patterns. The sentiment expressed by news outlet is often a reflection of how a population thinks or feels. This can be on a larger scale with national or international news editions, but also on a smaller regional scale with regional news that is specific to a smaller population, making the analysis of news on specific topics particularly interesting. Moreover, computer linguistic studies of this sort are rare due to a large majority of natural language processing studies being applied to the English language. France played a major role historically in how news agencies work and distribute news to this day, making it an ideal candidate for this study. Being home to the first news agency in the world, \emph{Agence France Presse}, was created under the reign of Louis Philippe I, the last king of France and second last monarch which was followed by the introduction of the law on the Freedom of the Press of 29 July 1881. This project correlates sentiment into quantitative metrics as a way to uncover potential relationships, this will be done through the extraction of sentiment proxies in news from a corpus of French news which was built as part of this project. 

\section{Motivation}\label{Motivation}
Looking back in history, it is clear that the world went and will go through many phases
whether it be a pandemic, a war or a revolution to name a few. What makes the recent COVID-19 pandemic particularly interesting is that it is the first to be spread throughout the world during the digital age, creating new behaviours on the web.   
In the context of a global pandemic in a digital age, the role of news agencies is more important than ever, with the world population in quarantine all looking for the same information, seeking to understand the situation they are in.

One could manually read newspaper articles from many different sources about a topic in an attempt to understand how information is conveyed to the population, however such a study would be limited in the manpower available and would require a strict, well defined approach to do so. Taking such an approach is thus limited in its nature.

Natural Language Processing (NLP) is a sub-field of computer science which enables analysis of text, Sentiment Analysis makes use of NLP in order to extract subjective information from texts. Examples of this include extracting and quantifying the polarity or certain emotions from text.
To this day, most of the studies performed and the applications of Sentiment Analysis have been focused on the English language, leaving space for many new studies, specifically in languages that are constructed significantly differently such as French.

The French media publishes vast amounts of news articles daily, all of which can be used to extract meaningful information such as to ultimately understand hidden patterns in how information is communicated to the population in the context of a global emergency that is the COVID-19 pandemic.
This thesis attempts to automate the process of discovering such hidden patterns through the use of sentiment analysis on French news.

\section{Research Question}\label{chap: Research Question}

The research question for this thesis can be summarized as follows:

\begin{center}
    \emph{To what extent does the sentiment of French news media relate to France's COVID-19 metrics?}
\end{center}

\section{Objectives}\label{chap: Research Objectives}
To structure the work done and address the research question effectively, the following key objectives were defined:
\begin{enumerate}
    \item Gather a large corpus containing high quality French news articles relevant to COVID-19
    \item Create specialised dictionaries targeted to measure pandemic-related sentiments
    \item Perform sentiment analysis on the corpus to retrieve measures for various sentiments and term frequencies
    \item Conduct statistical analysis on the results in order to identify and quantify potential relationships
\end{enumerate}

\section{Report Structure}\label{Report Structure}

The next chapter will start with a look into the current literature on the role of the media in the context of a global emergency such as the COVID-19 pandemic (Chapter \ref{Literature Review}). This is followed by a review and discussion of the current literature for sentiment analysis, which will cover the main approaches used for its various applications. This chapter concludes with a discourse of the statistical methods used throughout this report.

The literature review is followed by a section on the method used (Chapter \ref{Method}). The method includes details of the approach taken to solve the research objectives previously laid out. It is split into three parts: data gathering which focuses on constructing and collecting data for the analysis, sentiment analysis which details the actions performed on the corpus to measure various sentiments and finally a section on the statistical analysis performed.

The following section (Chapter \ref{Evalutation}) discusses the results obtained from applying the method as previously described. It will begin with a review of the built corpus and include some initial analysis. Following will be the results of the corpus through the use of sentiment analysis and various dictionaries, some of which were built specifically for this study. This chapter will close with a discussion of the statistical analysis performed on the obtained metrics.

Finally, this report concludes with a review of the method which was used to perform this study (Chapter \ref{Conclusion}). Then a commentary on the key findings and the possible improvements will be carried out which outlines directions for future work. The thesis will come to an end with final comments on the project, providing insights into the challenges that were encountered and a brief reflection on the personal achievements that were made throughout this project.
